---
title: "Photodegradation Data Analysis: Mass Loss"
author: "Robert Logan"
date: "March 19, 2018"
output: html_document
---

These data come from the 6, 12, 18, 30, and 36 month measurement times for the photodegradation experiment.

Racks deployed on: : 23 Jul 2016
t=6 removed on: 20 Jan 2017
t=12 removed on: 16 Jul 2017
t=18 removed on: 12 Jan 2018
t=30 removed on: 30 Jan 2019
t=36 removed on: 20 Jul 2019

# Load and organize data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list = ls())

library(ggplot2)
library(gridExtra)
library(tidyverse)
library(broom)
library(car)
library(lme4)
library(MuMIn)
library(lsmeans)
library(cowplot)


```

# Mass loss figure (Figure 2A)

```{r}

# Load data
Data_Photo <- read.csv("/Photodegradation Data LONG.csv", head=T)
Data_Photo$Treatment <- factor(Data_Photo$Treatment, levels = c("No Shelter", "UVpass", "UVblock", "Shade"))
Data_Photo$RackTiller <- paste(Data_Photo$Rack,Data_Photo$Stem)

# Convert appropriate columns to numeric
Data_Photo$Index <- as.numeric(Data_Photo$Index)
Data_Photo$Rack <- as.numeric(Data_Photo$Rack)
Data_Photo$Stem <- as.numeric(Data_Photo$Stem)
Data_Photo$StemAbsolute <- as.numeric(Data_Photo$StemAbsolute)
Data_Photo$Removal <- as.numeric(Data_Photo$Removal)
Data_Photo$Time <- as.numeric(Data_Photo$Time)
Data_Photo$Days <- as.numeric(Data_Photo$Days)
Data_Photo$InitialDryMass.g. <- as.numeric(Data_Photo$InitialDryMass.g.)
Data_Photo$Mass.g. <- as.numeric(Data_Photo$Mass.g.)

# Check
dim(Data_Photo)
names(Data_Photo)
head(Data_Photo)

# Calculate mass loss and mass loss percentages. Note: these aren't actually expressed as precentages but as proportions.
Data_Photo$MassLoss <- Data_Photo$Mass.g. - Data_Photo$InitialDryMass.g.
Data_Photo$MassLossPercent <- Data_Photo$MassLoss / Data_Photo$InitialDryMass.g.
Data_Photo$MassRemaining <- (1 + Data_Photo$MassLossPercent) * 100

# Remove faulty data
Data_Photo <- subset(Data_Photo, Data_Photo$Status=="good")
dim(Data_Photo)
Data_Photo <- subset(Data_Photo, Data_Photo$RackTiller != "10 3")
dim(Data_Photo)

# Generate figure
fig.MassLoss <- ggplot(data = Data_Photo, aes(x = Time, y = MassRemaining, col = Treatment)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 1) +
  stat_summary(fun.data = mean_se, geom = "line") +
  theme_bw() +
  scale_color_manual(values=c("gold", "orange", "red", "black")) +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  coord_cartesian(xlim = c(0, 36)) +
  xlab("Months") +
  ylab(bquote(atop("Mass Remaining", "(Percent)")))
fig.MassLoss



```

## Stand-alone mass loss figure

```{r}

# Add zeros for plotting
zeros <- expand_grid(Treatment = unique(Data_Photo$Treatment),
                     Removal = 0,
                     Days = 0,
                     Time = 0,
                     MassRemaining = 100)

# Generate figure
ggplot(data = Data_Photo %>%
         full_join(zeros),
       aes(x = Days, y = MassRemaining, col = Treatment)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 1) +
  stat_summary(fun.data = mean_se, geom = "line") +
  theme_bw(base_size = 20) +
  scale_color_manual(values=c("gold", "orange", "red", "black")) +
  coord_cartesian(ylim = c(85, 100)) +
  xlab("Days") +
  ylab(bquote(atop("Mass Remaining", "(Percent)")))

```


# Cuticle photos (Figure 2B)

```{r}

# Load and cleanup data
data.cuticle <- read.csv("/Microscope photo data.csv",header=T)

data.cuticle$Treatment <- factor(data.cuticle$Treatment, levels = c("No Shelter", "UVpass", "UVblock", "Shade"))
data.cuticle$RackTiller <- paste(data.cuticle$Rack,data.cuticle$Tiller)

data.primary <- subset(data.cuticle, data.cuticle$Type=="")
data.primary.f <- subset(data.primary, data.primary$Side=="F")
data.primary.b <- subset(data.primary, data.primary$Side=="B")

# Generate figure
fig.FrontPics <- ggplot(data = data.primary.f, aes(x = Time, y = Score, col = Treatment)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 1) +
  stat_summary(fun.data = mean_se, geom = "line") +
  theme_bw() +
  coord_cartesian(ylim=c(0.5,5.5), xlim = c(0, 36)) +
  scale_color_manual(values=c("gold", "orange", "red", "black")) +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  xlab("Months") +
  ylab(bquote(atop("Cuticle Damage", "Score")))
fig.FrontPics

```

# Moisture uptake figure (Figure 2C)

```{r}

# Load data
Data_Moisture <- read.csv("/MoistureUptake.csv", header=T)
Data_Moisture$Treatment <- factor(Data_Moisture$Treatment, levels = c("No Shelter", "UVpass", "UVblock", "Shade"))

# Calculate normalized water uptake (mg water / mm^2)
Data_Moisture$WaterPerSA <- (Data_Moisture$WaterMass / Data_Moisture$SA) * 1000
hist(Data_Moisture$WaterPerSA, breaks = 12)

# Moisture figure
fig.Moisture <- ggplot(data = Data_Moisture, aes(x = Time, y = WaterPerSA, col = Treatment)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 1) +
  stat_summary(fun.data = mean_se, geom = "line") +
  theme_bw() +
  scale_color_manual(values=c("gold", "orange", "red", "black")) +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)) +
  coord_cartesian(xlim = c(0, 36), ylim = c(0, 0.16)) +
  xlab("Months") +
  ylab(bquote(atop("Water Uptake", "(mg/mm" ^2 * ")")))
fig.Moisture

```

# Composite figure (Figure 2A-C)

```{r}

Grob.MassLoss <- ggplotGrob(fig.MassLoss)
Grob.FrontPics <- ggplotGrob(fig.FrontPics)
Grob.Moisture <- ggplotGrob(fig.Moisture)

plot_grid(Grob.MassLoss, Grob.FrontPics, Grob.Moisture,
          ncol=1,
          align="v")

# ggsave(
#   filename = "CompositeFigure1",
#   plot = last_plot(),
#   device = "eps",
#   path = "",
#   scale = 1,
#   width = 6,
#   height = 6,
#   units = "in",
#   dpi = 300,
#   limitsize = TRUE)

```


# Compare measurements taken across the three experiments (Figure 2D-E)

```{r}

Data_Moisture_simple <- Data_Moisture %>%
  select(Rack, Tiller, Treatment, UVtreatment, Time, WaterPerSA)

Data_Photo_simple <- Data_Photo %>%
  filter(Removal == Time) %>%
  select(Rack, Stem, Time, MassRemaining) %>%
  rename("Tiller" = "Stem")

dataMerged <- left_join(Data_Moisture_simple, Data_Photo_simple)

dataMerged2 <- left_join(dataMerged,
                         data.primary.f %>%
                           select(Rack, Tiller, Time, Score)) %>%
  mutate(MassLoss = 100 - MassRemaining)

dataMerged2$Treatment <- factor(dataMerged2$Treatment, levels = c("No Shelter", "UVpass", "UVblock", "Shade"))

# Cuticle damage vs. moisture uptake
ggplot(dataMerged2, aes(x = Score, y = WaterPerSA, col = Treatment)) +
  theme_bw(base_size = 20) +
  geom_jitter(size = 3, width = 0.25) +
  xlab("Cuticle Damage Score") +
  # ylab(expression("Water uptake (mg H"[2]* "O mm" ^ -2* ")")) +
  # ylab(expression("Water Uptake (mg H"[2]* "O/mm" ^ 2 * ")")) +
  scale_color_manual(values=c("gold", "orange", "red", "black")) +
  ylab(expression("Water Uptake (mg/mm" ^ 2 * ")")) +
  theme(legend.position = "none")
model1 <- wilcox.test(dataMerged2, formula = WaterPerSA ~ Score)
anova(model1)

# Moisture uptake vs. mass loss
ggplot(dataMerged2, aes(x = WaterPerSA, y = MassLoss, col = Treatment)) +
  theme_bw(base_size = 20) +
  geom_point(size = 3) +
  # xlab(expression("Water uptake (mg H"[2]* "O mm" ^ -2* ")")) +
  # xlab(expression("Water Uptake (mg H"[2]* "O/mm" ^ 2 * ")")) +
  scale_color_manual(values=c("gold", "orange", "red", "black")) +
  xlab(expression("Water Uptake (mg/mm" ^ 2 * ")")) +
  ylab("Mass Loss (%)") +
  theme(legend.position = "none")
# model2 <- lm(dataMerged2, formula = MassLoss ~ WaterPerSA * Treatment)
# anova(model2)

```



